---
title: "Trends in Educational Attainment: Results"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: word_document
---

```{r textPrelim}
opts_chunk$set(
echo=FALSE,results='hide',warning = FALSE,message = FALSE,error=FALSE,cache=FALSE
)
```

```{r textPrelim2, cache=TRUE}
library(knitr)
knit('overTime.Rmd')
knit('gapsOverTime.Rmd')
```

```{r textFuncs}
pfun <- function(strs)
  switch(strs,
    `***`="p<0.001",
    `**`="p<0.01",
    `*`="p<0.05",
    `.`="p<0.1")

years <- paste0('20',names(overTime)[1],' to 20',names(overTime)[length(overTime)])
```

```{r colnames}
for(i in 1:length(analysis)){
  if(NCOL(analysis[[i]]$trendStars)>1)
    colnames(analysis[[i]]$trendStars) <-
      if('deaf'%in%names(analysis[[i]])) colnames(analysis[[i]]$deaf$ests) else colnames(analysis[[i]]$ests)
}
```

```{r colnamesGaps}
for(i in 1:length(gaps)){
  if(length(gaps[[i]]$stars)>1)
    names(gaps[[i]]$stars) <- colnames(gaps[[i]]$ests)
}
```

# High School Attainment

From `r years`, high school
completion rates for deaf people grew by roughly
`r round(analysis$hsTot$ests['Growth',1])` percentage
points--a rate of `r round(analysis$hsTot$ests['Trend (Adj)',1],1)`
percentage points per year (PPY) (`r pfun(analysis$hsTot$trendStars[2])`)
(Table 1).
When the sample was restricted to deaf people ages 25-34, the
estimated trend over time was `r round(analysis$hsByAgeCat$deaf$ests['Trend (Adj)',1],1)` PPY
(`r pfun(analysis$hsByAgeCat$trendStars[2,1])`).

High school completion trends were similar for deaf males and females, at
 `r round(analysis$hsSex$deaf$ests['Trend (Adj)','Male'],1)`
(`r pfun(analysis$hsSex$trendStars[2,2])`), and
 `r round(analysis$hsSex$deaf$ests['Trend (Adj)','Female'],1)`
(`r pfun(analysis$hsSex$trendStars[2,1])`) PPY, respectively.
Trends were more pronounced for deaf Latnix
(`r round(analysis$hsRace$deaf$ests['Trend (Adj)','Hispanic'],1)` PPY;
 `r pfun(analysis$hsRace$trendStars[2,'Hispanic'])`) and deaf African American
 (`r round(analysis$hsRace$deaf$ests['Trend (Adj)','African American'],1)` PPY;
 `r pfun(analysis$hsRace$trendStars[2,'African American'])`) people than
for deaf white people
(`r round(analysis$hsRace$deaf$ests['Trend (Adj)','White'],1)` PPY;
 `r pfun(analysis$hsRace$trendStars[2,'White'])`).
 Estimated trends for other racial and ethnic groups were too imprecise to draw conclusions.
 Within racial and ethnic categories, trends in high school completion were similar between deaf females and deaf males, though we estimated a higher trend for deaf African American Females
(`r round(analysis$hsRaceF$deaf$ests['Trend (Adj)','African American'],1)` PPY;
 `r pfun(analysis$hsRaceF$trendStars[2,'African American'])`) than for males
 (`r round(analysis$hsRaceM$deaf$ests['Trend (Adj)','African American'],1)` PPY;
 `r pfun(analysis$hsRaceM$trendStars[2,'African American'])`).

## Trends in High School Attainment Gaps

From `r years` the age-adjusted hearing-deaf gaps in high school completion narrowed from
roughly `r round(gaps$hsTot$ests['2008',1],1)` in 2008 to `r round(gaps$hsTot$ests['2018',1],1)` in 2018, a trend of rougly `r round(gaps$hsTot$ests['trend',1],2)` PPY
(`r pfun(gaps$hsTot$stars)`) (the negative sign indicates a narrowing gap).

Curously, this acheivement gap trend is not apparent in the 25-34 age group; instead, it appears to have been driven largely by changes in the proportions of deaf, but not hearing, people ages 35-54 reporting high school diploma attainment.

The high school gap appeared to narrow more for women
(`r round(gaps$hsSex$ests['trend','Female'],2)` PPY; `r pfun(gaps$hsSex$stars['Female'])`)
than for men
(`r round(gaps$hsSex$ests['trend','Male'],2)` PPY; `r pfun(gaps$hsSex$stars['Male'])`)
though the difference in trends is not statistically significant.
Nevertheless, as of 2018 the high school completion gap remains higher for women, at
`r round(gaps$hsSex$ests['2018','Female'],2)` than for men `r round(gaps$hsSex$ests['2018','Male'],2)`.
There was little evidence of differeng trends between racial/ethnic groups, or within racial ethnic groups between genders.
That said, the estimated trend in the high school acheivement gap for African American Women was particularly pronounced--`r round(gaps$hsRaceF$ests['trend','African American'],2)` PPY (`r pfun(gaps$hsRaceF$stars['African American'])`).

# Associate's Degree Attainment

From `r years`, Associate's Degree
completion rates for deaf people grew by roughly
`r round(analysis$ccTot$ests['Growth',1])` percentage
points--a rate of `r round(analysis$ccTot$ests['Trend (Adj)',1],1)`
percentage points per year (PPY) (`r pfun(analysis$ccTot$trendStars[2])`)
(Table 1).
When the sample was restricted to deaf people ages 25-34, the
estimated trend over time was `r round(analysis$ccByAgeCat$deaf$ests['Trend (Adj)',1],1)` PPY
(`r pfun(analysis$ccByAgeCat$trendStars[2,1])`).

Associate's Degree completion trends were similar for deaf males and females, at
 `r round(analysis$ccSex$deaf$ests['Trend (Adj)','Male'],1)`
(`r pfun(analysis$ccSex$trendStars[2,2])`), and
 `r round(analysis$ccSex$deaf$ests['Trend (Adj)','Female'],1)`
(`r pfun(analysis$ccSex$trendStars[2,1])`) PPY, respectively.
Trends were similar for deaf Latnix
(`r round(analysis$ccRace$deaf$ests['Trend (Adj)','Hispanic'],1)` PPY;
 `r pfun(analysis$ccRace$trendStars[2,'Hispanic'])`), deaf African American
 (`r round(analysis$ccRace$deaf$ests['Trend (Adj)','African American'],1)` PPY;
 `r pfun(analysis$ccRace$trendStars[2,'African American'])`), and
 deaf white people
(`r round(analysis$ccRace$deaf$ests['Trend (Adj)','White'],1)` PPY;
 `r pfun(analysis$ccRace$trendStars[2,'White'])`).
 Estimated trends for other racial and ethnic groups were too imprecise to draw conclusions.
 Within racial and ethnic categories, trends in Associate's Degree completion were similar between deaf females and deaf males.

## Trends in Associate's Degree Attainment Gaps

From `r years` the age-adjusted hearing-deaf gaps in the proportion of people earning at least an Associate's Degree narrowed from
roughly `r round(gaps$ccTot$ests['2008',1],1)` in 2008 to `r round(gaps$ccTot$ests['2018',1],1)` in 2018, a trend of rougly `r round(gaps$ccTot$ests['trend',1],2)` PPY
(`r pfun(gaps$ccTot$stars)`).

This trend appears to be driven entirely by changes in in the deaf-hearing associate's degree gap for men, which narrowed by approximately `r round(gaps$ccSex$ests['trend','Male'],2)` PPY (`r pfun(gaps$ccSex$stars['Male'])`).
In contrast, the gap for women narrowed by
 `r round(gaps$ccSex$ests['trend','Female'],2)` PPY (p=`r round(gaps$ccSex$pvals['Female'],2)`).

Trends in the associate's degree gap for age or racial/ethnic categories were mostly too small or measured with too much noise to draw conclusions.
One exception is the trend in the gap for white males, which was
`r round(gaps$ccRaceM$ests['trend','White'],2)` PPY (`r pfun(gaps$ccRaceM$stars['White'])`), comparable to the estimated trends in gaps African American and Latinx males,
`r round(gaps$ccRaceM$ests['trend','African American'],2)` and `r round(gaps$ccRaceM$ests['trend','Hispanic'],2)` PPY, respectively, which, however, were not statistically significant due to imprecision.
Curiously, the gap for males in the "Other" racial/ethnic category appeared to _increase_ at a rate of
`r round(gaps$ccRaceM$ests['trend','Other'],2)` (`r pfun(gaps$ccRaceM$stars['Other'])`), though this is likely a spurious result.


# Bachelor's Degree Attainment

From `r years`, Bachelor's Degree
completion rates for deaf people grew by roughly
`r round(analysis$bachTot$ests['Growth',1])` percentage
points--a rate of `r round(analysis$bachTot$ests['Trend (Adj)',1],1)`
percentage points per year (PPY) (`r pfun(analysis$bachTot$trendStars[2])`)
(Table 1).
When the sample was restricted to deaf people ages 25-34, the
estimated trend over time was `r round(analysis$bachByAgeCat$deaf$ests['Trend (Adj)',1],1)` PPY
(`r pfun(analysis$bachByAgeCat$trendStars[2,1])`).

Bachelor's Degree completion trends were somewhat higher for deaf females than males, at
 `r round(analysis$bachSex$deaf$ests['Trend (Adj)','Female'],1)`
(`r pfun(analysis$bachSex$trendStars[2,'Female'])`), and
 `r round(analysis$bachSex$deaf$ests['Trend (Adj)','Male'],1)`
(`r pfun(analysis$bachSex$trendStars[2,'Male'])`) PPY, respectively.
Trends were similar for deaf Latnix
(`r round(analysis$bachRace$deaf$ests['Trend (Adj)','Hispanic'],1)` PPY;
 `r pfun(analysis$bachRace$trendStars[2,'Hispanic'])`), deaf African American
 (`r round(analysis$bachRace$deaf$ests['Trend (Adj)','African American'],1)` PPY;
 `r pfun(analysis$bachRace$trendStars[2,'African American'])`), and
 deaf white people
(`r round(analysis$bachRace$deaf$ests['Trend (Adj)','White'],1)` PPY;
 `r pfun(analysis$bachRace$trendStars[2,'White'])`).
 Estimated trends for other racial and ethnic groups were too imprecise to draw conclusions.
 Within racial and ethnic categories, trends in Bachelor's Degree completion were similar between deaf females and deaf males.

## Trends in Bachelor's Degree Attainment Gaps

The trend in the deaf-hearing gap in Bachelor's degree attainment was estimated as
`r round(gaps$bachTot$ests['trend'],2)` PPY; this estimate was statistically insignificant
(p=`r round(gaps$bachTot$pvals,2)`).
However, the BA gap for males did appear to narrow, at a rate of
`r round(gaps$bachSex$ests['trend','Male'],2)` PPY (`r pfun(gaps$bachSex$stars['Male'])`).
None of the other subgroups we studied exhibited a statistically significant trend in the bachelor's degree gap from `r years`.
